### 2019-08-02 Chlorophyll.Rmd
chloro

---
title: "Chlorophylla"
author: "Amieroh Abrahams"
date: "18 February 2019"
output: html_document
---

## Script info
The aim of this script is to extract the chlorophyll a netCDF files and convert it to CSV. The chlorophylla data were obtained from MODIS Aqua. The aim of working with these chlorophyll a data is to examine how the chlorophyll a concentration varies with upwelling.

```{r}
library(ncdf4)
library(data.table)
library(tidyverse)
library(reshape2)
library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
library(lubridate)
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(ggpubr)
library(fasttime)
library(FNN)
library(heatwaveR)
```

This function convert chlorophyll netCDF datasets to a single CSV datset

```{r}
MODIS_chlor.dir <- "/home/amieroh/Documents/Data/Datasets/Chlorophyll_a"
MODIS_chlor.csv.dir <- "/home/amieroh/Documents/Data/Datasets"

region <- "BC" # Benguela Current
coords <- bbox[, region]
coords <- c(-35, -20, 10, 20) # this is the BC

ncList <- list.files(path = MODIS_chlor.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
strt.date <- str_sub(basename(ncList[1]), start = 2, end = 8)
end.date <- str_sub(basename(ncList[length(ncList)]), start = 2, end = 8)
nc.init <- nc_open(ncList[1])
LatIdx <- which(nc.init$dim$lat$vals > coords[1] & nc.init$dim$lat$vals < coords[2])
LonIdx <- which(nc.init$dim$lon$vals > coords[3] & nc.init$dim$lon$vals < coords[4])
nc_close(nc.init)

ncFun <- function(nc.file = nc.files, csv.dir = csv.dir) {
nc <- nc_open(nc.file)
  instrument <- ncatt_get(nc, 0, "instrument")$value
  platform <- ncatt_get(nc, 0, "platform")$value
  product_name <- ncatt_get(nc, 0, "product_name")$value
  fNameStem <- substr(product_name, 17, 38)
  timeStamp <- substr(product_name, 2, 8)
  origin <- paste0(substr(timeStamp, 1, 4), "-01-01")
  date <- as.Date(as.numeric(substr(timeStamp, 5, 7)), origin)
  chl <- round(ncvar_get(nc,
                   varid = "chlor_a",
                   start = c(LonIdx[1], LatIdx[1]),
                   count = c(length(LonIdx), length(LatIdx))),
               3)
  dimnames(chl) <- list(lon = nc$dim$lon$vals[LonIdx],
                        lat =  nc$dim$lat$vals[LatIdx])
  nc_close(nc)
  chl <-
    as.data.table(melt(chl, value.name = "chl"), row.names = NULL) %>%
    mutate(t = date) %>%
    na.omit()
  fwrite(chl,
         file = paste(csv.dir, "/", region, "-", instrument, ".",platform, ".",
                      fNameStem, "-", strt.date, "-", end.date, ".csv", sep = ""),
         append = TRUE, col.names = FALSE)
  rm(chl)
}

llply(ncList, ncFun, csv.dir = MODIS_chlor.csv.dir, .parallel = TRUE)

# MODIS_Chloro <- "~/Documents/Masters_2019/MastersProject/Data"
# MODIS_Chloro <- fread(paste0(MODIS_Chloro, "/BC-MODIS.Aqua.L3m_8D_CHL_chlor_a_9km-2002185-2018345.csv"),
#             col.names = c("lon", "lat", "chloro", "date"))

# save(MODIS_Chloro, file = "Data/MODIS_Chloro.RData")
```

## Setting up the chlorophylla dataset

```{r}

# Loading the data
load("Data/MODIS_Chloro.RData") # Extracted chlorophyll data
load("Data/site_list_sub.RData") # Site list of all the sites along the west coast

chloro_plot <- MODIS_Chloro %>%
  filter(date == min(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = chloro))

unique_pixel <- MODIS_Chloro %>% 
  select(lon, lat) %>% 
  unique()

match_index <- knnx.index(data = as.matrix(unique_pixel[,1:2]),
                          query = as.matrix(site_list_sub[,5:6]), k = 1)

# Matching the sites within the chlorophyll data to the sites along the west coast
pixel_match <- unique_pixel[match_index,] %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  mutate(site = site_list_sub$site)

chloro_match <- MODIS_Chloro %>%
  unite(col = combi, lon, lat, sep = "/", remove = F) %>%
  filter(combi %in% pixel_match$combi)

chloro_match %>%
  filter(date == max(date)) %>%
  ggplot(aes(x = lon, y = lat)) +
  geom_raster(aes(fill = chloro))

combined_chloro_sites <- chloro_match %>%
  left_join(pixel_match, by = c("combi", "lon", "lat")) %>% 
  dplyr::mutate(date = as.Date(date))

# Save the complete dataset
# save(combined_chloro_sites, file = "Data/combined_chloro_sites.RData")

Chloro_plot_complete <- ggplot(combined_chloro_sites, aes(x = date, y = chloro)) +
  geom_line() +
  facet_wrap(~site, ncol = 1)
Chloro_plot_complete

# Plotting only years 2014-2015
daily_chloro_data_tester <- combined_chloro_sites %>% 
  filter(year(date) %in% seq(2014, 2015))

Chloro_ggplot(daily_chloro_data_tester, aes(x = date, y = chloro)) +
  geom_line() +
  facet_wrap(~site, ncol = 1)
Chloro_plot_filtered

# # With the chloro pixels filter, it is now faster to create the daily data
daily_chloro_data <-  %>%
  dplyr::mutate(date = as.Date(date)) %>%
  dplyr::group_by(combi, lon, lat, date) %>%
  dplyr::summarise(chloro = mean(chloro, na.rm = TRUE)) %>%
  dplyr::group_by(combi, lon, lat) %>%
  tidyr::nest() %>%
  dplyr::mutate(clims = purrr::map(data, ts2clm, x = date, y = chloro,
                     climatologyPeriod = c("2002-07-13", "2018-12-04"), 
                     maxPadLength = 8)) %>% 
  dplyr::select(-data) %>%
  tidyr::unnest() %>%
  select(-doy, -seas, -thresh)

## Plotting daily data
# ggplot(daily_chloro_data, aes(x = date, y = chloro)) +
#   geom_line() +
#   facet_wrap(~combi, ncol = 1)

# daily_chloro_match_sites <- daily_chloro_data %>%
# left_join(pixel_match, by = c("combi", "lon", "lat", "month", "day", "year"))
```