### 2019-09-17 Write_up/write_up.Rmd
Edits

---
title: "Thesis - Upwelling and MHW/Cold spells analyses"
author: "Amieroh Abrahams"
date: "30 January 2019"
output:
  word_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    highlight: zenburn
    toc: yes
  html_document:
    fig_caption: yes
    fig_retina: 2
    fig_width: 7
    highlight: espresso
    theme: sandstone
    toc_float: no
  bookdown::html_document2:
    fig_caption: yes
    fig_retina: 2
    fig_width: 7
    highlight: espresso
    theme: sandstone
    toc_float: no
  pdf_document:
    toc: yes
    toc_depth: 4
    highlight: default
    keep_tex: yes
    latex_engine: xelatex
fontsize: 10pt
geometry: margin=1in
language: Australian
mainfont: Minion Pro
monofont: Courier
papersize: A4
sansfont: Myriad Pro
tables: yes
bibliography: ../bib_master.bib
---

# Chapter 1: Introduction

## Eastern boundary upwelling systems

Globally, there are four major eastern boundary upwelling systems (EBUS). Upwelling systems include the Benguela Current off south-western Africa, the Canary Current off north-western Africa together with its northern extension off the Iberian Peninsula of south-western Europe, the Peru-Humboldt Current off western South America, and the California Current off the western continental USA and north-western Mexico [@bakun ; @pauly]. Each of these systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans. EBUS cover only 1% of the world’s ocean surface, but despite this these are among the most productive regions of the world [@pauly]. This is due to the  systems exposure to coastal upwelling, which is defined as the process whereby cold, nutrient rich, high concentrated CO2, low pH, and low oxygenated waters are pushed to the surface as a result of alongshore winds interacting with the earth’s rotation [@huyer]. As a result of upwelling, diverse marine fauna and flora are able to thrive in these areas due to the abundance of nutrients and favourable temperatures.

Upwelling is primarily caused by alongshore, equator ward winds. These winds are caused by cross-shore atmospheric pressure gradients, and these gradients occur predominantly during heating periods. Bakun proposed that if there was an increase in greenhouse gases, it would cause more heating during the day and more cooling during the night causing the temperature gradient to increase resulting in a strong pressure gradient, Ultimately effecting wind patterns and therefore upwelling. Several people have tried to model this using different datasets of SST and wind pattern data at various resolutions and initially found that increases in gases, such as CO^2 causes a reduction in trade winds, and therefore a reduction in upwelling. These models however, were based on coarse data and when higher resolutions were used, such as for the Benguela EBUS, the models have shown that upwelling will intensify in accordance to the upwelling intensification hypotheses [@hsieh; @mote; @lima].

### Ecology and primary production within upwelling zones

The duration and intensity of coastal upwelling are known to play a critical role in the phenology of key marine ecosystem processes. Similarly, changes in the characteristics of upwelling are shown to cause large disturbances to ecosystems at several trophic levels [@wang]. Cold, nutrient rich, low pH, and low oxygen upwelled waters fuels primary production as a consequence of the development of phytoplankton blooms [@ryther; @Cushing; @mcGillicuddy; @oschlies]. In these upwelling systems, the high productivity supports vast and diverse marine populations. Mesoscale and sub-mesoscale oceanic movements are key in managing the exchange and transportation of organic resources and marine organisms from the coastal zone to the open ocean. Most of the world’s major fishing industries are dependent on upwelling, as this process brings nutrient rich waters into the euphotic zone, where it forms the foundation of a localized food web that ecologically differs from surrounding, non-upwelled regions [@bosman; @barshis]. Whilst nutrients play a primary role in the survival of these systems, the cool temperatures of these waters are critical as well. There is an increasingly large body of research that has provided evidence showing that sea surface temperatures (SST) have and are continuing to increase, however some regions still show a lower or non-significant increase, and in some cases cooling trends are evident (REFERENCES). Upwelling occurring along the Cape south coast is proven to have less of an economic importance in contrast to the west coast. Inshore regions and embayments that are protected from wind action are likely to be subjected to intense episodes of primary production [@graham]. Vander Wouder et al. (2006) found that these sites represent the highest levels of chlorophyll concentration. Biotic systems within the Benguela upwelling system may be negatively influenced by the changes in oxygen levels given their adaption to natural variable conditions [@capone].

For several decades, regime shifts have become an important factor of marine research. Regime shifts are often defined as a shift from one stable ecosystem state to another. This process affects several species within a short period of time [@cury; @jarre; @blamey]. Regime shifts are thought to be driven by abiotic processes such as climate change [@polovina; @blamey]. The southern Benguela upwelling system supports a wide variety of commercially-exploited fish species and exhibits substantial variability in both oceanographic and biological components. Over the past  three decades, a number of species occurring within this region have undergone shifts in their distributions and Roy et al. (2007) associated these shifts to cooling waters along the inner shelf of the Agulhas bank on the south coast, east of Cape Agulhas [@blamey]. This coast is associated with high levels of harmful toxic algal blooms, which negatively impact commercial and recreational interests along the coastal region. These harmful blooms are often attributed to flagellated species with the red tide representing dinoflagellate blooms (Pitcher et al. 1993A, @pitcher1998]. Formations of red tides are often related to the prevailing winds of the southern Benguela which govern most hydrodynamic processes on the continental shelf. Red tide events are greatest during the latter part of the upwelling season during transitions in the synoptic weather patterns which result in diminished upwelling activity and increased thermal stratification (Pitcher et al. 1993a).

### Atmosphere

Throughout regions surrounding the Benguela EBUS, the atmosphere is dominated by anticyclonic high-pressure cells with quasi-stationary positions. The South Atlantic Ocean High is situated along the west, drawing cool, dry air onto the west of the subcontinent [@van]. Solar heating during summer may result in the development of low-pressure cells, known as heat lows, which are absent during the winter [@tyson]. As the anticyclones shift north during winter months, the cold westerlies have a substantial impact on the weather of the southern tip of the South African subcontinent [@van]. The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings].

## Sea surface temperature and data collection

Water temperature affects the survival, abundance, distribution and growth of marine species (REFERENCE). As such, a clear understanding of the relationship between marine ecosystems and their thermal tolerance is necessary for conservation and management planning. Future predictions in biogeographical patterns and ecosystem functioning may also be estimated based on various climatic behaviours. Regardless of the objectives, whether it be studies based on climate change, extreme events or ecosystem functioning, it is important that the research is conducted through accurate, reliable datasets. Unfortunately, not all data are created equally and under the same conditions and as such, not all datasets are appropriate for a particular task.

### *In situ* data

Having coastal, * in situ* seawater temperature data is a rare privilege to researchers. Most biological and ecological research requires coastal temperature data. The South African Coastal Temperature Network (SACTN). The availability of the *in situ* collected coastal temperature data products for several decades provides a reliable source of accurate coastal seawater temperature data [@smit]. SACTN has collected coastal seawater temperature along the South African coastline from as early as 1972, with contributions from a total of seven different organisations and governmental departments. At present, this dataset contains temperatures of 129 sites along the three distinct coasts of the South African coastline. Temperature recordings are still ongoing at many of these sites, but unfortunately, the durations and extents of these are uneven per site. Data collected for this region started on 13 September 1972 and concluded on 26 January 2017, with recordings still continuing daily. During the 1970s, a total of 11 time series began recording. A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s.

The uneven and inconsistent nature of the SATCN dataset can be attributed to several factors. For instance, as a result of variances in the equipment used by the various organisations and governmental bodies who recorded the data. One problematic outcome of this sampling methodology was that, due to specific work of the South African Weather Service (SAWS) and Kwa-Zulu Natal Sharks Board (KZNSB), two thirds of the SACTN data that were sampled were obtained with hand-held thermometers and recorded at a data precision of 0.500 ◦ C, however, the current global standard is to make use of underwater temperature recorders (UTRs), which have a data precision of 0.001◦ C [@jarraud]. Due of these inconsistencies, this dataset required verification from an independent source before its use in research. Smit et al. (2013) thoroughly compared the temperature data within the SACTN dataset to those of several remotely-sensed SST products and found bias results. Using these *in situ* collected temperature data along with remotely sensed SST data requires a clear understanding of the characteristics of the contributing data sources in order to understand how accurate, reliable and useful these long-term measurements are for use in climate change studies.

It should also be acknowledged that most of the time series within the SACTN dataset do not have meta-data records that meet the international climate change research requirements [@aguilar]. Specifically, uncertainty lies within the drift of the instrument, who was sampling and where the exact sampling locations were collected throughout the collection period. There is certainty that all of the thermometer time series were collected by thermometers and that the UTR time series were collected by UTRs in this study.  Given these issues associated with the mete-data, recent work has been started to develop a nation wide standard.

For this analysis, the seawater temperature data, obtained from the seven different sources, were combined and formatted into standardized comma delineated value (CSV) files which allowed for a fixed methodology to be used across the entire dataset. Prior to data analysis, all data points exceeding 35°C and/or below 0°C were removed as these were considered as outliers. These data points were then changed to NA (not available) so as to not interfere with analysis. All analyses were conducted in R software version 3.4.2 (http://www.r-project.org/). The data used within this study and comprehensive script used for data analyses, and production of figures can be found at https://github.com/AmierohAbrahams/MastersProject.

## Satellite data

Reliable datasets are often difficult to maintain and obtain. In the field of biological science research, it is often abiotic data products that are impaired thus leading to the use of satellite derived SST products for the study of specific fields such as coastal and marine ecology as well as oceanography. These datasets however, may be inaccurate along the coast and local *in situ* collected data should be used when possible (Smit et al., 2013). Remotely sensed data products are obtained by several different organisations worldwide. Satellites launched into space detect the infra-red light emitted by the ocean which is then taken as a proxy from SST. Given the extensive distance at which these satellites record data, a certain level of coarseness is displayed within the dataset. This is demonstrated as the size of the pixels one sees in the final product.

Satellite SST products undergo a range of quality control and validation processes (e.g. Reynolds and Smith, 1994; Brown et al., 1999; Martin et al., 2012), making the data reliable and more useful. As previously mentioned, satellites may show a warm or cold bias along the coastal region but are still useful for broader scale investigations (Smit et al., 2013, Smale and Wernberg, 2009; Castillo and Lima, 2010; Dufois and Rouault). Remotely-sensed SST products are obtained at different resolutions, and as a result do not represent the same level of bias against *in situ* collected seawater temperature. Research done by Smit et al (2013) showed that MODIS represented the least amount of bias.

This study made use of various satellite-derived Sea Surface Temperature (SST) datasets. One of datasets used was the National Oceanic and Atmospheric Administration (NOAA) ¼° daily Optimum Interpolation Sea Surface Temperature (OISST). This dataset is an analysis of seawater temperature collected by different platforms (satellites, ships and buoys) on a global grid. The gridded daily SST product assimilates both remotely sensed and **in situ** sources of data to create a level-4 gap free product (Banzon et al., 2016). A spatially complete sea surface temperature map is produced by interpolation to fill in gaps within the dataset. The method used to fill these gaps include the adjustment of satellite and ship observations which compensate for bias differences (NOAA.GOV, https://www.ncdc.noaa.gov/oisst). There are two types of daily OISST, named after the relevant satellite sensors, these are the Advanced Very High Resolution Radiometer (AVHRR) and Advanced Microwave Scanning Radiometer on the Earth Observing System (AMSR-E). Currently the AVHRR has the longest record starting from 1981 to present day. Infrared instruments, like AVHRR, can make observations at relatively high resolutions. Microwave instruments, such as AMSR-E, can measure SSTs in most weather conditions but not adjacent to land. The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region. The AVHRR datasets have provided global SSTs for more than four decades [@reynolds; @Pearce; @corlett].

The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, used as the second dataset, was produced using satellite instruments with datasets spanning from1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/). MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates, roughly at 1 km intervals. This dataset is currently among the highest resolution SST dataset available. The goal in creating the MUR SST is to develop a coherent and consistent daily map of SST at the highest spatial resolution possible. The MUR sea surface temperature dataset combines data from the Advanced Very High Resolution Radiometer (AVHRR), Moderate Imaging Spectroradiometer (MODIS) Terra and Aqua, and Advanced Microwave Spectroradiometer-EOS (AMSR-E) instruments in an optimal way in order to produce 1km global maps representing sea surface temperature.

The Moderate-resolution Imaging Spectroradiometer (MODIS) is a scientific instrument (radiometer) on board the NASA Terra and Aqua satellite platforms with a spatial resolution in 0.041° longitude-latitude coordinates. MODIS was launched in 1991 and 2002 to study global dynamics of the earth’s atmosphere, ice and ocean. MODIS captures data in 36 spectral bands ranging in wavelength from 0.4 um to 14.4 um and at varying spatial resolutions (2 bands at 250 m, 5 bands at 500 m and 29 bands at 1 km), providing complete global coverage of the Earth every one to two days [@hosoda; @gentemann, @dong]. MODIS Terra Global level 3 mapped thermal SST products consists of sea surface temperature data derived from the 11 and 12um thermal IR infrared (IR) Bands (NASA.GOV, http://oceancolor.gsfc.nasa.gov).

The CMC dataset comprises the fourth dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10 km resolution constructed by the Canadian Meteorological Center (CMC). The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with *in situ* observations of SST from ships and buoys from the ICOADS program.

The fifth dataset, the Global 1 km SST (G1SST) analysis, representing a 0.01° resolution, uses satellite data from sensors that include the Advanced Very High Resolution Radiometer (AVHRR), the Advanced Along Track Scanning Radiometer (AATSR), the Spinning Enhanced Visible and Infrared Imager (SEVIRI), the Advanced Microwave Scanning Radiometer-EOS (AMSRE), the Tropical Rainfall Measuring Mission Microwave Imager (TMI), the Moderate Resolution Imaging Spectroradiometer (MODIS), the Geostationary Operational Environmental Satellite (GOES) Imager, the Multi-Functional Transport Satellite 1R (MTSAT-1R) radiometer, and *in situ* data from drifting and moored buoys [@dash]. This dataset produces a product that combines satellite data with the model output (PO.DAAC, https://podaac.jpl.nasa.gov/).

### Climate change

Climate change as a result of anthropogenic activities is a global concern of both scientific and political importance. As the human population grows, the rate of global warming increases, resulting in profound effects on marine and other ecosystems. Climbing temperatures create a host of additional changes to marine systems, such as rising sea levels, increased ocean stratification, decreased sea-ice extent, and altered patterns of ocean circulation and precipitation [@doney]. Direct effects on ocean temperature changes and chemistry may alter the functioning and behaviour of marine organisms and in turn, these changes lead to a change in species integration and trophic pathways.

Significant research studies accept that climate change and variability have great impacts on marine ecosystems [@stenseth; @harley; @hoegh]. The Intergovernmental Panel on Climate Change (IPCC) predicts future projections with a mean increase in air temperature of 1.1°C by the 2020’s and 1.8°C by the 2040’s as a result of the increase of greenhouse gas concentrations [@solomon]. A large amount of this heat will be absorbed by the world’s ocean. In current years, as a result of changes in air temperature, the heat content of the ocean has increased significantly. Given ocean warming in association with their physical changes such as sea level rise and declines in sea ice, marine life is expected to intensify in future decades. Global warming represents distinct warming and cooling periods throughout the 20th and 21st century (IPCC, 2007). Global warming however is not evenly distributed across the global ocean, and thus several regions are warming faster or slower than the global average [@levitus; @santos]. In particular, the Atlantic Ocean contributes most to the increase of heat content [@nerem; @levitus].

As a result of climate change, research studies provide significant evidence relating to the increase in the warming of the ocean surface. This warming slowly propagates down through the water column and in turn increases thermal stratification [@gruber]. Strong stratification limits the depth at which water is upwelled and hence controls the amount of nutrients brought up [@chhak; @jacox2011; @jocox2016]. The interplay between upwelling winds and stratification determines the amount of nutrients present within the system and is therefore strongly linked to biological productivity. Globally, as a result of climate change, stratification of the water-column between 0 and 200 m has increased by an estimated 4% between 1971 to 2010 [@stocker]. Increased ocean stratification as a result of surface warming may reduce the ventilation of deep-water masses. As the deep water is being upwelled to the surface, decreases in oxygen and pH levels are predicted with an increase in nutrient levels [@rykaczewski].  Generally, regional ocean climate oscillations like the Benguela Niño for the Benguela Current System impose considerable variability on temperatures and complicate attempts to assess long-term trends [@demarcq]. It is predicted that these climate oscillations may increase the variance with climate change [@timmermann; @kuzmina; @sydeman] which may result in unusual trends.

As previously mentioned, upwelling winds in EBUSs are generated by pressure differences between continental thermal lows and oceanic highs. As warming intensifies, pressure gradients between land and sea increases, resulting in a greater intensity of upwelling winds [@bakun2010]. In rare cases, projected upwelling increases may overcome the countervailing effects of upper-ocean warming and stratification to cause regional cooling [@auad]. Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes [@keeling]. Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material [@hofmann; @doney].

Climate change has severe ecological consequences on both plants and animals (Bradley et al. 1999; Walther et al. 2002). In the past two decades, sea surface temperature has risen in many regions and this seasonal rise in temperature can act as an important cue in species survival, reproduction and abundance (Cushing 1982; Cushing 1990). The different responses of species as a result of climate change are important to understand as these changes affect, the trophic dynamics and fisheries yield of several regional ecosystems among many things (Murawski 1993). Understanding the relationship between global warming and climate change as a result of anthropogenic activity is a vast and challenging process. In a bid to further this understanding, Rockström et al. (2009) introduced the concept of planetary boundaries. According to this concept, within each of these boundaries exists a level of safety that should not be exceeded. According to this concept, the planetary boundary for biodiversity loss, has been surpassed [@steffen]. Similarly, the safety boundary for warming of the globe by greenhouse gas emission has also been surpassed. The idea of planetary boundaries is essential in that it helps in understanding the threats humans cause via the continued destruction of the natural environment. Consequently, studies of anthropogenic influences on the ecosystem are generally divided into land or sea. Terrestrial ecosystems are more familiar to the general populace; whereas marine ecosystems are associated with high levels of primary production [@Chavez] and are highly at risk as a result of climate change [@osterblom]. Climate change has been proven to decrease primary production thus affecting trophic levels and leading to trophic turbulence and the destruction of marine fauna [@lewandowska; @dell, @mcCauley].

Climate change affects coastal species such as oysters, corals and, sea grass. Lord (2017) also found evidence that suggested that a change in climate as a result of rising sea temperatures may increase the spread of invasive species and increase the chances of an outbreak of diseases [@altizer]. Research by Mead et al (2013) found that coastal ecosystems are at high risk as a result of climate change, later, Whitfield et al (2016) confirmed that coastal bio-diversity has been observed to be affected. Given the importance of primary productivity, it was determined that the Benguela upwelling system within the South African region should become the focus of this thesis. Here the increase in global temperatures is one of the main focus points along with chlorophyll concentration. By making use of these two measurements, an analysis of how chlorophyll concentration varies with temperature changes could be conducted.

## Benguela

The area of study for this thesis will focus on the west coast of South Africa, which is an eastern boundary upwelling system. The Benguela EBUS is unique in that it is bounded at both the equatorward and poleward ends by warm water regimes. These currents are forced locally by the wind stress field off southwest Africa [@nelson ; @fennel]. This upwelling system stretches along the south-western coast of Africa from southern Angola to Cape Agulhas [@cole]. Water is coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms]. The main upwelling period in the Benguela, north of 25°S extends mainly from March to November, peaking during August [@shannon1985], but upwelling persists throughout the year. Upwelling season in the northern Benguela extends between September and March, with very little upwelling taking place during winter (June-August). In the central Benguela and also at Cape Frio, upwelling is continuous throughout the year [@shannon1985]. The cold Benguela waters are bound in the south by the warm Agulhas retroflexion region and in the north by the southward flowing Angola current. The flow of the Benguela is often considered to be topographically guided [@nelson; @barange; @barange]. The Benguela system is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole].

Upwelling in the northern Benguela is in phase with the seasonal insulation cycle, whereas in the southern Benguela it is not. This results in a pronounced seasonal cycle in SST in the north while in the south there is no clear SST cycle.  The seasonal warming occurring during the late summer season off northern and central Namibia is a regular occurrence and is associated with warmer, higher salinity water of Angolan origin into the region [@o; @shannon]. There have been a number of warm and cool periods documented in the northern and southern Benguela [e.g. @stander; @walker; @shannon1984; @boyd; @brainard]. Walter (1937) found a major climatic event in the northern Benguela region of anonymously warm water. This unusual warm water may have resulted due to reduced upwelling or an increase southward advection of warm water. Another warming event occurred during the summer and autumn months of 1984 [@shannon1986]. The presence of this anomalous warm water during 1984 followed two cooler than normal years in the northern Benguela. Shannon et al. (1986) illustrated the warm (21-23°C) water near Walvis Bay as a large warm water patch offshore, west of Luderitz. The absence of upwelling between Walvis Bay and north of Luderitz is evident, however, the upwelling centre in the Benguela system [@shannon1986], and upwelling was minimal. Interestingly, in the southern Benguela region, there was no evidence of any unusual patterns in temperature distributions. The seasonal temperature pattern within the northern Benguela was somewhat similar to that in the eastern equatorial Atlantic and the regular southward intrusion of warm water into the northern Benguela during the summer season. During these warming periods the equatorward wind stress over the northern Benguela region was stronger than normal; yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region.

Northward, the Benguela Current Upwelling system is linked by the Angola Benguela Front which converge between the Benguela current and the Benguela coastal current as well as the warm saline waters of the Angola Current [@peterson]. Research done by Yamagata and Iizuka (1995) discovered that coastal Kelvin waves, originating from equatorial Kelvin waves, carry the downwelling process southward. Southward, the coastal upwelling system is bounded by the Agulhas Bank. South of the African coast the warm Agulhas Current, western boundary current, flows in from the Indian Ocean and returns eastwards while generating warm eddies that spin off in a north-westerly direction in the Benguela current [@duncombe]. The region surrounding the Benguela experiences a persistent along shore wind. This wind is associated with the St. Helena high pressure system. Lutjeharms and Meeuwis, (1987) also found a strong relationship between the intensity of upwelling events and the strength and direction of prevailing winds. These favourable upwelling equatorial winds represent a maximum at 25°S and decrease towards the northern and southern boundaries of the Benguela upwelling system. Continental shelf bathymetry and upwelling favourable winds provide a large-scale upwelling mechanism in the southern Benguela, whereas local topography and meteorology create an alternating pattern of active and passive upwelling circulations along the coast. Winds occurring within the southern regions represent a seasonal variation, reaching a maximum during spring and summer months [@boyd; @shannon1996; @fennel]. Winds occurring within this region tend to show increasing patterns away from the coast whereas, winds occurring in the northern region exhibits relatively less seasonal variation [@shannon1985; @pitcher]. Upwelling occurring along the south coast of the Cape Province within South Africa, specifically along the coastline is generated by local wind, with wind increasing somewhat away from the coast. This facilitated the idea of a wind stress curl, as discussed in Bakun and Nelson (1991) and Shannon and Nelson (1996).

Benguela Niños are defined as anomalous warm events occurring between the southward flowing Angola current and the Benguela upwelling system off south western Africa [@shannon1986]. These warm events profoundly impact fisheries and the climate of this region. They are proven to induce unexpected rainfall events and drastically influence fish abundance and distribution [@rouault; @boyer]. Evidently, Benguela Niños are expressed as regions of abnormal, persistent high sea surface temperatures. Many researchers often associate the equatorial interannual variability pattern in the Atlantic to the ENSO phenomenon found within the Pacific region [@sutton; @zebiak]. Benguela Niños have been recorded to occur in 1934, 1949, 1963, 1984 and 1995 [@shannon; @gammelsrod]. Picaut (1985) suggested that eastern tropic oceans are governed by remote wind stress effects through equatorial wave dynamics. Anomalies in trade winds may result in Kelvin waves that propagate eastward along the equator, inducing a deepening or a lifting of the thermocline. It is possible that equatorial Kelvin waves may play a role in the evolution of Benguela Niños. ***Need to make changes to this paragraph, random but feels important***

### Extreme events

Extreme thermal events have a range of broad scale definitions, but they can each be classified into two categories, namely, cold-spells [e.g. @gunter; @lirman; @boucek] or heatwaves [e.g. @gordon; @stott; @perkins]. These events occur rapidly, and unexpectedly and have significant impacts on species diversity, distributions, growth and survival where they take place. These events are often undetected, but studies have found that extreme events have occurred as recently as approximately four decades ago. Jentsch et al. (2007) were the first to discuss this mode of climate change and brought it to greater focus. Steps have since been taken towards understanding the drivers, impacts and consequences of these events. Subsequently, research has predicted an increase in the occurrence of these extreme events as a result of long-term warming of the planet via the process of global warming [@mccabe; @harley; @schoepf].

Little is known about extreme events occurring within the South African region. Few South African researchers have specifically focussed on investigating this topic, and published literature is sparse. During the 1980’s however, Birkett and Cook analysed the impacts of extreme warming events on the life cycle of white muscles, *Donax serra Röding*. Further studies on extreme events occurred within the Florida Reef Tract, where anomolously cold events resulted in the mass mortality of coral reefs and fish. *Acropora cervicornis*, was found to be particularly susceptible to low temperatures and experienced a high mortality rate [@shinn; @davis]. While these phenomena are not as common as warming events, they have a dramatic and long-lasting effect on coral communities. Studies in January 2010 revealed that the Florida Reef Tract experienced a cold-water anomaly, marine cold spells (MCSs) with temperatures below 16 ◦ C recorded for up to 14 days [@lirman]. Species population distribution are often set by cold distribution boundaries.  More recently, Schlegel et al, (2017, 2018) examined the relationship between the nearshore and offshore occurrences of marine heatwaves and cold-spells as well as the influence of predominant atmospheric and oceanic patterns during Coastal Marine Heatwaves. It is now becoming accepted that extreme events are a threat to several ecosystems [@thompson]. In a warming planet one would assume that the occurrence and intensity of the severity of marine cold spells (MCSs) are likely to decrease. This thesis aims to address that assumption.

### Event detection

As specified earlier, the extreme cold events occurring along the west coast of South Africa, which is central to this thesis will be detected and correlated to upwelling events with the algorithm developed by Hobday et al. (2016). The events that recorded *in situ* data from SACTN would be compared against those detected in satellite derived SST. For this analyses, the R packages ‘RmarineHeatWaves’ and ‘heatwaveR’ were used to calculate the extreme events. In order to obtain the basis for these calculations, the temperature metrics must be aggregated into mean values based on the year in which they occurred. With these annual values it is possible to calculate long-term trends in extreme events.

## Analyses

The analyses conducted in this thesis varied according to the investigations performed. The one common thread uniting all of the analyses was that they were conducted in R-programming language (R Core Team, 2017). This provided reproducibility for all of the analyses and visualisations used.

## Problem statement:

With global warming continuing at a rapid pace, the amount of greenhouse gases within the atmosphere will subsequently increase in the future. This increase will undoubtedly have an impact on the world’s oceans and the marine life within it, both directly and indirectly. The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving the upwelling process within EBUS regions. Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora as well as human activities such as fishing. To assess changes in upwelling events, the aim of this investigation is to make use of SST and wind data to monitor the durations and intensities of upwelling events at different distances from the coastline within the Benguela EBUS.

Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015). TThereafter,Inorder to determine whether upwelling events had resulted in an increase in the onset of marine cold spells, all four easter boundry upwelling systems were considred.

## Aims and approach

The warming of the global climate as a result of anthropogenic activity, has negatively influenced both terrestrial and marine ecosystems, and ultimately inhibits humanity and ecological functioning [@mcCauley]. This climate variability, manifested as increases in the frequency and duration of extreme events is regarded as a great threat [@thompsonet]. Very few studies have addressed the occurrence and the changes in the intensity or duration of cold temperature spells or cold events within the Benguela region. This investigation examines recent trends and variation in the frequency and duration of upwelling events occurring in this region In this research project, various satellite derived observations of sea surface temperature (SST) and wind will be used to further investigate the Benguela upwelling system along the South African coastline, whilst considering cold events and warm events occurring within this region. Thereafter, satellite data will be used to examine upwelling on a global scale and whether these events may have influence the rate of onset of marine cold spells.

## Study 1

The primary aim of the first study was to make use of various remotely sensed SST and in-situ collected coastal seawater temperature data. The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset (Smit et al. 2013). It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and if these patterns that exist are evident within the various datasets.

## Study 2

The primary aim of the second study was to make use of remotely sensed satellite sea surface temperature for all of the four EBUS to determine whether or not a change exsit in the rate of onset of marine coldspells and marine heatwaves. It was also useful to examine how the intensity and frequency of upwelling events varied between datasets and over the years given the influence of climate change and sea surface temperature rises. Wind data used in these analyses were predicted to show clear patterns between weak and favourable upwelling winds and its impact on the intensity and duration of upwelling. The testing of the hypothesis would be made possible with a series of analyses that allows one to select specific time periods from a range of datasets. Thereafter analyses rate of onset of MCS was done.This research could also assists in the understanding of local oceanographic patterns along the west coast of South Africa and the impact anthropogenically forced climate change could have for the recent changes in patterns and long-term trends

## Research assumptions

The central assumption of this project was that remotely-sensed SST data and *in situ* seawater temperature data may be compared. Remotely sensed SST data and *in situ* seawater temperature data were obtained at different resolutions and so slight changes may be found within the different datasets. An additional assumption within this research focused on the influence of the changes of upwelling on marine cold spells. Very little research has been done on the driving factors leading to MCSs and which processes may result in an increase in the intensity and duration of these events. For the purpose of this study the influence of upwelling on MCSs was the primary focus.

# Paper 1
## Chapter 2 Upwelling

## Abstract

### Introduction

Upwelling is a physical oceanographic process in which cold, nutrient rich waters are pushed to the surface as a result of the motion caused by alongshore equatorial winds. These upwelling events are known to have a large influence on the abundance, diversity, and distribution of marine organisms, and moreover, critically affect marine ecosystem processes (Add in references). Globally, upwelling occurs throughout the world’s oceans, however, four major eastern boundary upwelling systems (EBUS) are recognised, each significantly impacting coastal systems. These respective systems are characterized as vast regions of coastal ocean occurring along the western shores of continents bordering the Pacific and Atlantic Oceans [@bakun; @pauly]. The oceanography of the inshore waters off the west coast of South Africa is dominated by an upwelling system which, along with the Peru, California and Humboldt currents as Eastern boundary upwelling regions [@andrews]. Eastern boundary upwelling regions only cover 1% of the world’s ocean surface but are among the most productive regions of the world. As such, there has been a constant interest in upwelling systems. Major upwelling programs have already been carried through off of Peru (e.g. WALSH. KELLEY, DUGDALE and FROST, 1971) and North-West Africa (e.g. BARBER, 1977).

Over the past three decades, climate change as a result of anthropogenic activities has established global concern and is shown to have negatively affected marine and terrestrial ecosystems. The consequences of these changes are also shown to have direct effects on the behaviour and functioning of marine organisms, thus ultimately affecting species trophic levels and integration [@doney]. According to the Bakun hypothesis, an increase in greenhouse gases will result in an increase in day-time warming and night-time cooling and ultimately cause an increase in temperature gradients which will then form stronger pressure gradients. Pressure gradients are known to affect wind patterns which will subsequently affect the intensity and duration of upwelling events [@hsieh; @mote; @lima].

Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references). These temperatures have increased as a result of anthropogenically driven climate change [@gruber]. Warming is shown to slowly move down into the water column thus increasing thermal stratification. This thermal stratification restricts the depth at which water is upwelled and therfore limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016]. Climate change impacting oceanic processes has major ecological consequences (Bradley et al. 1999; Walther et al. 2002). This is demonstrated in several regions increases in sea surface temperatures have been proven to greatly influence marine species survival, reproduction and abundance. Climate change has also been shown to largely influence coastal species such as corals, muscles and oysters. Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper. Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well (Murawski 1993) therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions.

The 3000km South African coastline, is bordered by the Agulhas current along the east coast and the EBUS Benguela current, along the west coast. These coastal regions display noticeable differences in seawater temperatures in comparison to each other, primarily due to the influences of the neighbouring ocean currents (Bolton et al. 2004; Mead et al. 2013; Schlegel and Smit 2016). The Benguela Current is driven by the anticyclone high pressure systems and eastward moving cyclones which determine the boundaries of the Benguela upwelling region (Shannon 2006; Hutchings et al. 2009). The atmospheric temperatures along the coastline are largely influenced by the Benguela Current [@van], however, upwelling systems may be forced by various atmospheric conditions [@Lutjeharms; @roberts; @hutchings]. Regionally, the Benguela Current assists in transporting cold water northwards from the Southern Ocean to the coast (Lüning 1990; Lutjeharms, Cooper, and Roberts 2000; Hutchings et al. 2009; Schlegel et al. 2017). Changes in the intensity and duration of upwelling events have proven to cause large changes in ecosystem functioning [@wang].

The Benguela system is bounded by the equatorward and poleward ends by warm water regimes. These currents are known to be forced by local wind stress fields off southwest Africa [@nelson; @fennel]. This particular EBUS stretches along the south-western coast of Africa, from southern Angola to Cape Agulhas [@cole], with water temperatures coldest and most persistent off the Namibian coast at 25°S and weakest at 17°S [@Lutjeharms]. The upwelling period in the Benguela north of 25°C persists throughout the entire year but peaks during August. The Benguela upwelling region is split into northern and southern systems by a zone of intense perennial upwelling activity in the Luderitz region within the Namibian region [@shannon; @cole]. Several studies have shown that extreme warming and cooling periods occurring in the northern and southern Benguela systems [@stander; @walker; @shannon1984; @boyd; @brainard]. Furthermore, Walter (1937) found that warm periods of global ocean temperatures may have resulted as a lack of upwelling or poor upwelling events. During the periods of warming, the equatorward wind stress over the northern region of the Benguela system was stronger than usual yet upwelling was suppressed by the southward movement of warm saline waters from the Angolan region. In the southern Benguela region, continental shelf bathymetry and strong upwelling favourable winds are known to result in more intense upwelling events. Upwelling occurring along the south coastline of the Southern Africa region are usually generated by local winds. Event-scale upwelling has been demonstrated to have substantial impacts on fisheries [@roy; @van].

The Cape peninsula is influenced by south-east trade winds during summer and westerlies during winter, thus a pronounced seasonal cycle is evident within patterns of wind movement across the coastline. During spring, summer, and autumn seasons, the anticyclonic high-pressure cell periodically ridges eastwards south of the continent. During this process, south easterly winds predominate, with peak frequencies occurring during spring and summer months resulting in more upwelling events during summer and winter months. During the winter seasons, low pressure cells are driven from the west to the east as a result of the atmospheric circumpolar waves and so north-westerly winds dominate during winter. South easterly winds dominate during summer but are still present for short durations during winter [@andrews]. Within a range of kilometres off the shore, wind conditions vary due to the effects of mountain funnelling within the region.

The dynamics of upwelling within the nearshore (<400m from the coastline) remains poorly understood. Upwelling metrics are often calculated far offshore because measurements are typically taken from large ships that cannot move close to the shore [@andrews]. Cold temperatures are important in describing species population distribution limits and are proven to set limits in the north or southwards higher latitudes (Firth et al., 2011). There has not yet been any understanding of how the upwelling process varies at local versus broader scales. In this study, it was hypothesize that upwelling is more dominant within coastal regions instead of broader scale. Very few studies have addressed the changes in the intensity or duration of cold temperature spells or cold events at varying distances from the coast within the Benguela region.

Strong temperature gradients existing between offshore and inshore oceanic waters may result in different upwelling event readings. Upwelling relatively close to the shore has been found to depend on the strength and direction of the wind and tends to vary based on seasonal wind variation [@andrews]. Inshore temperatures tend to be lower during summer months due to upwelling, but there is an increase in offshore temperatures in the later summer months as a result of warming of the South East Atlantic Surface waters[@andrews,@Shannon]. Due to the temperature gradients that exist inshore and offshore, this study focused on observing upwelling patterns from the coastal region to 50km from the coastline.

The central assumption of this project was that remotely-sensed SST data and *in situ* seawater temperature data may be compared. Remotely sensed SST data and *in situ* seawater temperature data were obtained at different resolutions and so slight changes may be found within the different datasets. The primary aim of the first study was to make use of various remotely sensed SST and in-situ collected coastal seawater temperature data. The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset [@smit]. It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and if these patterns that exist are evident within the various datasets. This investigation examined recent trends and variation of upwelling events at different distances from the coastline.

## Methods

## Site selection

At a broad scale, the South African coastline exhibits a large variation in seawater temperature and is divided into four bioregions, each with contrasting temperatures [@smit, @mead]. The western region of the coastline is dominated by the Benguela Current forming an Eastern Boundry Upwelling System (EBUS) [@hutchings] which provides a natural labratory for this study. Annual mean coastal seawater temperature within this region represents a coastal seawater temperature range of 12.3 ± 1.2°C at the weastern limit near the Namibian border. Seasonal upwelling is controlled by south-easterly trade winds and as a result reflect a distinct temperature variations representing much lower temperatures within the upwelling cells over a fairly narrow continental shelf found from the Cape Peninsula to Cape Columbine. This region represents monthly mean temperatures at generally intermediate at cold and warm temperate [@luning].  In order to accurately examine upwelling patterns at various distances from the coast several sites were selected as this provides a more accurate representation, the appropriate sites along the west coast of South Africa (i.e. sites influenced by the Benguela current) were selected. Since temperature data were not evenly recorded for each of the sites, the entire dataset was narrowed to only those sites with a time series of longer than 30 years. This yield a total of five sites within the Benguela region (Fig 1).

```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline. The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=8, fig.width=12, fig.pos="H"}

knitr::include_graphics("../Figures/site_map.png")
```

## Temperature data sources

The South African Coastal Temperature Network (SACTN) data serves as the *in situ* temperature dataset used in this study. This dataset consists of coastal seawater temperatures obtained from 129 sites along the South African coastline. The SACTN has collected SST data along the unique coastline of South Africa from as early as 1972 till present date, with contributions made from several organisations and govermental departments. As a result, temperature within this datasets were obtained by using different methods and instruments, such as hand held thermometers or Underwater temperature reorders( UTRs). The precision of the instruments used vary accordingly, with hand-held thermometers recording temperature at a precision of 0.5°C as opposed to the underwater temperature recorders which have an instrument precision of 0.001°C. The duration and extent of recordings per sites were uneven with the longest time series in the dataset being that of Gordons Bay which started on 13 September 1972 and concluded on 26 January 2017.  During the 1970s, a total of 11 time series began recording. A further 53 entries were added during the 1980s, 34 entries were added during the 1990s, and 18 entries were added during the 2000s. Recordings are still ongoing at many of these sites.

This study also made use of a range of remotely sensed temperature datasets. Satellites provide a system of large spatial and temporal data retrieval. Remotely sensed SST products are produced by a number of organisations across the globe. The AVHRR-only Optimally-Interpolated Sea Surface Temperature (OISST) was used to determine SST within the study region. The AVHRR datasets have been provided global SSTs for more than four decades (Reynolds and Smith, 1994; Pearce et al 2006). OISST is a global 1/4° gridded daily SST product that assimilates both remotely sensed and in situ sources of data to create a level-4 gap free product (Banzon et al., 2016). The Multi-scale Ultra-high Resolution (MUR) Sea Surface Temperature Analysis, the second dataset, is produced using satellite instruments with datasets spanning 1 June 2002 to present times (PO.DAAC, https://podaac.jpl.nasa.gov/). MUR provides SST data at a spatial resolution of 0.01° in longitude-latitude coordinates and is currently among the highest resolution SST datasets available. The CMC dataset constitutes the third dataset and is a version 3.0 Group for High Resolution Sea Surface Temperature (GHRSST) Level 4 dataset with a 10km resolution constructed by the Canadian Meteorological Center (CMC). The CMC dataset combines infrared satellite SST at numerous points in the time series from the AVHRR, the European Meteorological Operational-A (METOP-A) and Operational-B (METOP-B) platforms, and microwave SST data from the Advanced Microwave Scanning Radiometer 2 in conjunction with in situ observations of SST from ships and buoys from the ICOADS program.

An obvious previledge exist when making using of *in situ* data over satellite data as they provide a more accurate representation of the thermal properties closer to the coast where satellite temperatures do not capture temperature properties well and thus can more accurately explain upwelling within the coastal inshore environment. Smit et al (2013) have shown that satellite data collected along the South African coastline represent a warm biases reaching as high as 6°C over *in situ* temperatures within the nearshore environment.

Inorder to create a time series of each of the remotely sensed SST data and also to compare whether the same upwelling patterns exist at the broad-scale against the *in situ* time series, shore-normal transects were extended from the *in situ* sites a long the west coast of South Africa. Temperature values from the various satellite datasets were extracted at multiple distances from the coast. These distances are 15 x 15 km, 25 X 25 km and 50 X 50 km pixels along the transect (Fig. 1). The daily sea surface temperatures obtained from the various remotely sensed datasets within each of the transects were initiated inorder to produce offshore, broad-scale time series that could accurately be used as a dataset to be compared against the *in situ* time series.

## Wind data

Wind action was an important variable in this study as it was hypothesised that wind direction and wind speed may have a direct influence on upwelling; consequently, wind action was investigated for its impact on upwelling at specific sites along the west coast of South Africa. Wind data was obtained from the South African Weather Service (SAWS), and were provided at three-hour resolutions. Specific wind characteristics were measured, namely, wind direction (dirw) and wind speed (spw). Since the wind data were modelled at three-hour resolutions, they were converted into daily data points in order to compare them with the temperature data. The circular() function in R software was then used to create circular objects around the wave data in order to calculate the daily wind parameters. ERA5 hourly data on pressure levels was used as a second wind source. This dataset comprise of u and v variables from 1973 to present.

With temperature and wind values now corresponding to their respective sites and dates, the hypothesis testing upwelling patterns at different distances from the coastline could now be tested. To do this, it was important to identify upwelling.

# Defining and determining upwelling

Inorder to determine the change in upwelling signals at different distances from the coastline and to test which dataset shows clear, visible representation of these events, it was first necessary to define when upwelling was occuring. We then aimed to determine a threshold value for which we could identify when upwelling was occuring within this region. Given that upwelling is primarily caused by alongshore, equator ward winds, it was necessary to consider both temperature and wind values. To determine the upwelling condition, an upwelling index was calculated using the formula presented in Fielding and Davis (1989). The upwelling index provided an indication of the intensity of the upwelling process taking place using the following equation:

                  $$ UpwellingIndex = μ{(Cosθ − coastal angle)} $$

In this expression, μ represents the wind speed (m/s) and θ represents the wind direction which is measured in degrees. The coastal angle is the angle of each of the sites perpendicular to the coastline. The index relies heavily on wind speed and direction data inorder to determine the intensity of upwelling likely to occur. Thereafter, the 'exceedance' function within the heatwaveR package was used to determine the consecutive number of days at or above which the upwelling index value is ment to be. Here it was specified that the minimum duration for upwelling is 1 day with a threshold value obtained from the upwelling index as 1. Inorder to obtain a more approved threshold value, a percentile value was determined. To determine changes within the upwelling values, the 'detect event' fuction within the heatwaveR package was used. Thereafter, to determine the upwelling event metrics ( frequency and intensity) the detect event function was used as a first filter with a threshold of 25th percentile. Values obtained from the upwelling indeces were used as a second threshold. An R script (https://github.com/AmierohAbrahams/Upwelling_MCS) was used to calculate the individual upwelling events for both the *in situ* and remotely sensed SST data. Because upwelling events were calculated relative to percentile exceedences, rather than an absolute definition such as periods with temperatures above a fixed temperature threshold, upwelling events occured any time of the year with upwelling shown to be more dominant during spring and summer months. For the use of the heatwaveR package it was best to use a time series of 30years.