### 2019-09-18 Write_up/Abstract_edits/Paper1_abstract.Rmd
update

---
title: "Abstract"
author: "Amieroh Abrahams"
date: "18 September 2019"
output: html_document
---

## Abstract