### 2019-10-08 SST_patterns.Rmd
Plotting Bits

---
title: "SST_patterns"
author: "Amieroh Abrahams"
date: "19 August 2019"
output: html_document
---

# Are the same upwelling signals present at different distances from the coastline and throughout the different datasets

```{r}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE 
)

library(tidyverse)
library(lubridate)
library(ggpubr)
library(zoo)
library(FNN)
library(scales)
library(gridExtra)
library(circular)
library(fossil)
library(mapproj)
# library(ncdf4) # This was used to process NetCDF files in an earlier version
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)
library(ncdf4) # library for processing netCDFs
library(data.table)
# library(plyr) # RWS: Never load the plyr package as it interferes with the tidyverse
library(heatwaveR)
```

# Loading in all the data created using the code bellow

```{r}
load("Data/site_list_sub.Rdata")
load("Data/SACTN_US.RData")
load("Data/site_pixels.RData") # 5 decimal places
load("Data/OISST.RData") # 2 decimal places
OISST <- BC_avhrr_only_v2_Document_Document 
rm(BC_avhrr_only_v2_Document_Document ); gc()
load("Data/CMC.RData") # 1decimal places
```

```{r}
# SACTN_US <- SACTN_US %>% 
#   dplyr::rename(in_situ_temp = temp)
# 
# # Visualising the data
# temp_plot <- function(df){
#   plot <- ggplot(data = df, aes(x = date, y = in_situ_temp, colour = site)) +
#     geom_line(aes(group = site)) +
#     labs(x = "", y = "Temperature (°C)") +
#     theme(axis.text.x = element_text(angle = 45)) +
#     theme(legend.position = "top")
# }
# 
# SACTN_plot <- temp_plot(df = SACTN_US)
# SACTN_plot

# 25th quantile
```

Different distances from the coastline

```{r}
# load("Data/site_list_sub.Rdata")
# # xtable(site_list_sub, auto = TRUE)
# west <- site_list_sub
# west$coast <- "west" # Chnages wc to west
# 
# load("Data/africa_coast.RData")
# 
# ## Downloading the bathy data from NOAA
# # Download mid-res bathymetry data
# # sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
# # sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
# # colnames(sa_bathy) <- c("lon", "lat", "depth")
# # sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
# # save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")
# 
# # Loading in the newly downloaded bathymetry data               
# load("Data/bathy/sa_bathy.RData")

# # This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
# shore.normal.transect <- function(site, width = 2){
#   # Find the site on the coastline and it's nearest neighbour points
#   coords <- data.frame(lon = site$lon, lat = site$lat)
#   coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
#   coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
#   coords3 <- coords3[2:1,1:3]
#   # Define the shore normal transect bearing
#   heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
#   if(heading >= 360){
#     heading <- heading-360
#   } else {
#     heading <- heading
#   }
#   heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
#   return(heading2)
# }
# 
# # Creating the transects
# site_transects <- data.frame()
# for(i in 1:length(west$site)){
#  site <- west[i,]
#  site_transect <- shore.normal.transect(site, 2)
#  site_transects <- rbind(site_transects, site_transect)
# }
# 
# # Manually correcting Sea Point and Kommetjie
# site_transects$heading[4:5] <- 290 
# # save(site_transects, file = "Data/site_transects.RData")
# load("Data/site_transects.RData")
# 
# # This function takes one site (e.g. one set of lon/lats) and calculates a shore norm./subal transect
# # It then extracts a lat/ lon point every X kilometres until reaching a specified isobath
# 
# transect.pixel <- function(site, distances){
#   # Extract coordinates
#   coords <- data.frame(lon = site$lon, lat = site$lat)
#   # Find lon/ lats every X metres 
#   pixels <- data.frame()
#   # deep <- 999
#   # distance_multiplier <- 1
#   # while(deep > isobath){
#   for(i in 1:length(distances)){
#     coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
#     sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
#     bathy2 <- sa_bathy[sitesIdx,]
#     bathy2 <- bathy2[complete.cases(bathy2$depth),]
#     bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
#                          heading = site$heading, 
#                          distance = distances[i])
#     pixels <- rbind(pixels, bathy3)
#     coords <- coords2
#   }
#   if(nrow(pixels) < 1){
#     pixels <- data.frame(site, depth = NA)
#   }else{
#     pixels <- pixels
#   }+

#   return(pixels)
# }
# 
# # Pixel points
# site_pixels <- data.frame()
# for(i in 1:length(west$site)){
#   site <- site_transects[i,]
#   site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
#   site_pixels <- rbind(site_pixels, site_pixel)
# }
# 
# # Bounding box
#   # Only one is made in order to know how large the the geom_point() squares should be made to match
# bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
#                    xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
#                    ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
#                    ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])
# 
# # Determining the temperature at the various distances from the coast
# 
# # save(site_pixels, file = "Data/site_pixels.RData")
# load("Data/site_pixels.RData")

```

# Extrcting the MUR data

```{r}
# library(stringr)
# library(tidyverse)
# library(reshape2)
# library(ncdf4) # library for processing netCDFs
# library(plyr)
# library(lubridate)
# library(data.table)
# library(doMC); doMC::registerDoMC(cores = 7)
# 
# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
# 
# #          1         2         3         4
# # 12345678901234567890123456789012345678901
# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
# 
# read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
#   endDate <- str_sub(ncLast, start = 1, end = 8)
# 
# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
# 
#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
#     nc <- nc_open(ncFile)
#     pathLen <- nchar(paste0(ncDir, "/")) + 1
    # fNameStem <-
    #   substr(basename(ncFile), 10, 38)
#     fDate <- substr(basename(ncFile), 1, 8)
#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
#       round(4)
#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
#                           lat = nc$dim$lat$vals)
#     nc_close(nc)
#     sst <- as_tibble(melt(sst, value.name = "temp"))
#     sst$t <- ymd(fDate)
#     na.omit(sst)
#     fwrite(sst,
#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
#            append = TRUE, col.names = FALSE)
#     rm(sst)
#   }
# 
# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
# 
# 
# MUR <- read_csv("/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR/BC-JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
# names(MUR)<-c("lon","lat", "temp", "date")
# MUR <- MUR %>% 
#   mutate(temp = temp - 273.15)
# 
# JPL_L4UHfnd_GLOB_v01_fv04_MUR_20020601_20140727 <- read_csv("~/Documents/JPL-L4UHfnd-GLOB-v01-fv04-MUR-20020601-20140727.csv")
# 
# ######################################### SUBSET VIA REGION #################
# 
# # bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
# library(stringr)
# library(tidyverse)
# library(reshape2)
# library(ncdf4) # library for processing netCDFs
# library(plyr)
# library(lubridate)
# library(data.table)
# library(doMC); doMC::registerDoMC(cores = 7)
# ncDir <- "/home/amieroh/Documents/Data/Datasets/MUR/daily"
# csvDir <- "/home/amieroh/Documents/Data/Datasets/MUR/Extracted_MUR"
# #          1         2         3         4
# # 12345678901234567890123456789012345678901
# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
# 
# # ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
#   ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
#     coords <- bbox[, region]
#     nc <- nc_open(ncFile)
#     pathLen <- nchar(paste0(ncDir, "/")) + 1
#     fNameStem <-
#       substr(basename(ncFile), 10, 38)
#     fDate <- substr(basename(ncFile), 1, 8)
#     LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
#     LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
#       round(4)
#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
#                           lat = nc$dim$lat$vals)
#     nc_close(nc)
#     sst <- as_tibble(melt(sst, value.name = "temp"))
#     sst$t <- ymd(fDate)
#     na.omit(sst)
#     fwrite(sst,
#            file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
#            append = TRUE, col.names = FALSE)
#     rm(sst)
#   }
#   
#   
#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
#   endDate <- str_sub(ncLast, start = 1, end = 8)  
#   
# llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
```

# Extracting the CMC data

```{r}
# ncDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC"
# csvDir <- "/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted"
# 
# #          1         2         3         4         5         6
# # 123456789012345678901234567890123456789012345678901234567890
# # 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc
# read_nc <- function(ncDir = ncDir, csvDir = csvDir) 
#   ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
#   ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
#   ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
#   strtDate <- str_sub(ncFirst, start = 1, end = 8)
#   endDate <- str_sub(ncLast, start = 1, end = 8)
# 
# # ncFile <- '/home/amieroh/Documents/Data/Datasets/CMC/CMC_BC/20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc'
# 
#   ncFun <- function(ncFile = ncFile, csvDir = csvDir) {
#     nc <- nc_open(ncFile)
#     pathLen <- nchar(paste0(ncDir, "/")) + 1
#     fNameStem <-
#       substr(basename(ncFile), 10, 58)
#     fDate <- substr(basename(ncFile), 1, 8)
#     sst <- ncvar_get(nc, varid = "analysed_sst") %>%
#       round(4)
#     dimnames(sst) <- list(lon = nc$dim$lon$vals,
#                           lat = nc$dim$lat$vals)
#     nc_close(nc)
#     sst <- as_tibble(melt(sst, value.name = "temp"))
#     sst$t <- ymd(fDate)
#     na.omit(sst)
#     fwrite(sst,
#            file = paste0(csvDir, "/", fNameStem, "-", strtDate, "-", endDate, ".csv"),
#            append = TRUE, col.names = FALSE)
#     rm(sst)
#   }
# 
# llply(ncList, ncFun, csvDir = csvDir, .parallel = TRUE)
# 
# 
# CMC <- read_csv("/home/amieroh/Documents/Data/Datasets/CMC/CMC_extracted/Benguela_current/20000-CMC-L4_GHRSST-SSTfnd-CMC0.2deg-GLOB-v02.0-f-19910901-20170317.csv")
# names(CMC)<-c("lon","lat", "temp", "date")
# CMC <- CMC %>% 
#   mutate(temp = temp - 273.15)
# 
# # save(CMC, file = "Data/CMC.RData")
```

# Extracting the OISST data

```{r}
# bbox <- data.frame(BC = c(-35, -25, 15, 20), # Benguela Current
#                    CC = c(25, 35, 340, 355), # Canary Current
#                    CalC = c(35, 45, 225, 240), # California Current
#                    HC = c(-17.5, -7.5, 275, 290), # Humboldt Current
#                    row.names = c("latmin", "latmax", "lonmin", "lonmax"))
# 
# OISST.dir <- "/home/amieroh/Documents/Data/Datasets/OISSTv2/daily/netCDF/avhrr-only"
# OISST.csv.dir <- "/home/amieroh/Documents/Data/Datasets/OISST_subset"
# 
# #          1         2
# # 1234567890123456789012345
# # avhrr-only-v2.19810901.nc
# 
# # function to extract the dims and data from OISST netCDFs
# read_nc <- function(ncFile, region = region, csvDir = csvDir) {
#   coords <- bbox[, region]
#   nc <- nc_open(ncFile)
#   pathLen <- nchar(OISST.dir) + 1 # to account for the "/" that needs to be inserted
#   fNameStem <-
#     substr(ncFile, pathLen + 1, pathLen + 13)
#   fDate <- substr(ncFile, pathLen + 15, pathLen + 22)
#   LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
#   LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
#   sst <- ncvar_get(nc,
#                    varid = "sst",
#                    start = c(LonIdx[1], LatIdx[1], 1, 1),
#                    count = c(length(LonIdx), length(LatIdx), 1, 1)) %>%
#     round(4)
#   dimnames(sst) <- list(lon = nc$dim$lon$vals[LonIdx],
#                         lat = nc$dim$lat$vals[LatIdx])
#   nc_close(nc)
#   sst <-
#     as.data.table(melt(sst, value.name = "temp"), row.names = NULL) %>%
#     mutate(t = ymd(fDate)) %>%
#     na.omit()
#   fwrite(sst,
#          file = paste(csvDir, "/", region, "-", fNameStem, ".", strtDate, "-", endDate, ".csv", sep = ""),
#          append = TRUE, col.names = FALSE)
#   rm(sst)
# }
# 
# # the list of files
# ncList <- list.files(path = OISST.dir, pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
# strtDate <- str_sub(ncList[1], start = 15, end = 22)
# endDate <- str_sub(ncList[length(ncList)], start = 15, end = 22)
# 
# # apply the function
# system.time(llply(ncList, read_nc, region = "BC", csvDir = OISST.csv.dir, .parallel = TRUE))
# # system.time(llply(ncList, read_nc, region = "CC", csvDir = OISST.csv.dir, .parallel = TRUE))
# # system.time(llply(ncList, read_nc, region = "CalC", csvDir = OISST.csv.dir, .parallel = TRUE))
# # system.time(llply(ncList, read_nc, region = "HC", csvDir = OISST.csv.dir, .parallel = TRUE))
# 
# # Loading the data
# BC_avhrr_only_v2_Document_Document <- read_csv("~/Documents/OISST_subset/BC-avhrr-only-v2.Document-Document.csv" )
# names(BC_avhrr_only_v2_Document_Document)<-c("lon","lat", "temp", "date")
# 
# # Saving the data
# save(BC_avhrr_only_v2_Document_Document, file = "Data/OISST.RData")
```

# Using the upwelling metrics created in `upwell.IDX.Rmd` Identify when upwelling occurs at the particular site. 
# Is this upwelling event seen throughout the different distances from the coastline
# Using CMC, MUR, OISST and SACTN_US to determine wether the upwelling events detected are present in each of the datasets

## Function
Determining the temperatures at the various distances from the coastline

```{r}
# # These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
# # MUR <- MUR %>% 
# #   select(lon, lat) %>% 
# #   mutate(product = "MUR")
# 
# # Decided to work with OISST and CMC as both has a time series of 30years
# OISST_prod <- OISST %>% 
#   select(lon, lat) %>% 
#   unique() %>% 
#   mutate(product = "OISST")
# 
# CMC_prod <- CMC %>% 
#   select(lon, lat) %>%
#   unique() %>% 
#   mutate(product = "CMC")
# 
# # sat_data <- rbind(CMC_prod, OISST_prod) %>% 
#   #rbind(., CMC) %>% 
#   # select(product, lon, lat)
# 
# # sat_pixels <- sat_data %>% 
#   # select(product, lon, lat) %>%
#   # unique()
# 
# ## For testing the nest/map pipeline
# # df <- site_pixels %>%
# #   filter(site == "Lamberts Bay") %>%
# #   select(-site)
# 
# match_func <- function(df){
#   df <- df %>%
#     dplyr::rename(lon_site = lon, lat_site = lat)
#   OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
#                                                  as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
#     cbind(., df)
#   CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
#                                              as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
#     cbind(., df)
#   res <- rbind(OISST_index, CMC_index)
#   return(res)
# }
# 
# # Find the nearest pixels to each spot along the transect
# # NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
# # points to not fall inside of the same 25 KM pixel
# pixel_match <- site_pixels %>% 
#   group_by(site) %>% 
#   group_modify(~match_func(.x))
# 
# # You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
# OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
# CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))
# #SACTN_fill <- right_join(SACTN, filter(pixel_match, product == "SACTN"), by = c("lon", "lat"))
# 
# # Clean up some RAM space
# rm(OISST, CMC); gc()
# 
# # sites <- c("Port Nolloth", "Lamberts Bay", "Saldanha Bay", "Sea Point")
# # # In the SACTN dataset Hout Bay only has data until 2005. Hout Bay will now be removed from this study
# # # Hout Bay will be ignored
# selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")
# 
# OISST_fill <- OISST_fill %>%
#   filter(site %in% selected_sites)
# 
# CMC_fill <- CMC_fill %>%
#   filter(site %in% selected_sites)
# 
# 
# # All the code below can be viewed in upwell_IDX.Rmd
# # Next to run upwelling index
# # UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
# load("Data/UI_angle.RData")
# 
# upwelling <- UI_angle %>% 
#   dplyr::rename(temp = ui.saws) %>%
#   group_by(site) %>%
#   # mutate(min_t = min(t), 
#   #        max_t = max(t)) %>% 
#   nest() %>% # apply the following functions to all of the variables in te dataset
#   mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
#          # NB: A threshold of 3 appeared to be far to strict
#          # purr::map - apllies a function to each element of a vector
#          exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
#   # Detect consecutive days in exceedance of a given threshold.
#   # mutate() %>% 
#   select(-data, -clim) %>% 
#   unnest() %>%
#   filter(row_number() %% 2 == 1) %>%
#   unnest() %>% # creates a column for each variables
#   dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
#   select(site, t, ui.saws, exceedance) 
# 
# # Now applying the Upwelling func
# # NB: This only pulls out the event results and not the climatology results
# # This is done to keep the output tidy because group_modify() may only create data.frame type outputs
# # To create a list output one would use group_map(), 
# # but this then loses the labels of which sites etc. the results belong to
# detect_event_custom <- function(df){
#   res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
#   return(res)
# }
# 
# ts2clm_custom <- function(df){
#   # The climatology base period used here is up for debate...
#   # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
#   res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
#   return(res)
# }
# 
# # Calculate the upwelling event metrics
# upwelling_detect_event <- function(df){
#   upwell_base <- df %>% 
#     dplyr::rename(t = date) %>% 
#     group_by(site, product, heading, distance, lon, lat) %>% 
#     group_modify(~ts2clm_custom(.x)) %>% 
#     left_join(upwelling, by = c("site", "t")) %>%
#     filter(!is.na(exceedance)) %>%
#     group_by(site, product, heading, distance, lon, lat) %>% 
#     group_modify(~detect_event_custom(.x))
#   }
# 
# OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
# save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
# CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
# save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
# 
# # Here we remove the site Hout Bay so that we have a long time series. The length of Hout Bay time series ends in 200. Many sites change from here
# SACTN_US <- SACTN_US %>%
#   filter(site %in% selected_sites)
# 
# load("Data/site_list_v4.2.RData")
# 
# SACTN_upwell_base <- SACTN_US %>%
#   left_join(site_list[,c(4, 5, 6)], by = "index")%>% 
#     dplyr::rename(t = date) %>% 
#     group_by(site, lon, lat) %>% 
#     group_modify(~ts2clm_custom(.x)) %>% 
#     left_join(upwelling, by = c("site", "t")) %>%
#     filter(!is.na(exceedance)) %>%
#     group_by(site, lon, lat) %>% 
#     group_modify(~detect_event_custom(.x))
# 
# save(SACTN_upwell_base, file = "Data/SACTN_upwell_base.RData")
```

# PLotting events

Loading the data
```{r}
load("Data/OISST_upwell_base.RData")
load("Data/CMC_upwell_base.RData")
load("Data/SACTN_upwell_base.RData")
```

To accurately compare 
  Start date CMC start date 1991-09-02 end date 2017-03-15
  Start date OISST start date 1981-09- 01  and end date 2018-09-28
  Start date SACTN start date 1973 and end date 2017-12-16

```{r}

lm_coeff <- function(df){
  res <- lm(formula = val ~ date_peak, data = df)
  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
}
# Changes in upwelling metrics
lm_func <- function(df){
  upwell_lm <- df %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var, distance) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
}

OISST_lm <- lm_func(df = OISST_upwell_base)
CMC_lm <- lm_func(df = CMC_upwell_base)

SACTN_lm <- SACTN_upwell_base %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
```

As a common date we will use a start date from 1992-01-31 and end date 16-12-31

```{r}
# Generalosed linear models

# Generalized Linear Models- Look at the difference in metrics of in situ upwelling and remotely sensed SST data upwelling 
# Generalized linear models are fit using the glm( ) function. The form of the glm function is glm(formula, family=familytype(link=linkfunction), data=)

# trends_duration_func <- function(df){
#   upwellmodel <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$duration[SACTN_upwell_base$event_no])), 
#                      family = poisson(link = "log"))
#   upwellmodel0 <- glm(SACTN_upwell_base$duration[SACTN_upwell_base$event_no] ~ 1, family = poisson(link = "log")) # intercept only
#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's pseudo-R2
#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
# }
# 
# OISST_trends_duration_func <- trends_duration_func(df = OISST_upwell_base)
# CMC_trends_duration_func <- trends_duration_func(df = CMC_upwell_base)
# SACTN_trends_duration_func <- trends_duration_func(df = SACTN_upwell_base)
# 
# 
# ### Intensity: Chnaged the family and link because of the negative values within the dataset 
# trends_intensity_func <- function(df){
#   upwellmodel <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ seq(1:length(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no])), 
#                      family = gaussian(link = "identity"))
#   upwellmodel0 <- glm(SACTN_upwell_base$intensity_mean[SACTN_upwell_base$event_no] ~ 1, family = gaussian(link = "identity")) # intercept only
#   dat2 <- data.frame(site = SACTN_upwell_base$site, upwelltrend = round(as.numeric(coef(upwellmodel)[2]*10),1),
#                        upwellR2 = round(1-logLik(upwellmodel)/logLik(upwellmodel0),2), # McFadden's psesudo-R2
#                        upwell.val = round(coef(summary(upwellmodel))[,4][2],2))
# }
# 
# OISST_trends_intensity_func <- trends_intensity_func(df = OISST_upwell_base)
# CMC_trends_intensity_func <- trends_intensity_func(df = CMC_upwell_base)
# SACTN_trends_intensity_func <- trends_intensity_func(df = SACTN_upwell_base)
```

# From the years 1992 - 2017

```{r}
# Which pixels showed the highest upwelling counts within the different datasets
# Which site showed the most intense upwelling within the different datasets

metric_func <- function(df){
 metrics <-df %>% 
  filter(year(date_start) %in% 1992:2017) %>% 
  #mutate(year = year(date_start)) %>% 
  group_by(distance, site) %>% 
  summarise(mean_intensity = mean(intensity_mean),
            sum_events = sum(event_no)) %>% 
   select(site, distance, mean_intensity, sum_events)
}

OISST_metrics <- metric_func(OISST_upwell_base)
CMC_metrics <- metric_func(CMC_upwell_base)

# No of u pwelling events found at the different sites
SACTN_metrics <- SACTN_upwell_base %>% 
  filter(year(date_start) %in% 1992:2017) %>% 
  group_by(site) %>% 
  summarise(mean_intensity = mean(intensity_mean),
            sum_events = sum(event_no)) %>% 
  select(site, mean_intensity, sum_events)

# Create a table showing these metrics

```

# Wind plots

```{r, fig.cap = "Windrose diagram representing the wind direction and speed for each of the sites", fig.height = 10, fig.width = 15}
library(ggradar)
library(dplyr)
library(scales)
library(tibble)
load("Data/UI_angle.RData")

# Detects NA values, no NA values are present
# Plotting wind variables
source("Functions/wind.rose.R") # This it may be related to the countmax = NA in the function

wind_daily_renamed <- UI_angle %>% 
  mutate(dir_circ = ifelse(dir_circ < 0, dir_circ+360, dir_circ)) %>% 
  dplyr::rename(spd = mean_speed) %>%
  dplyr::rename(dir = dir_circ) %>% 
  filter(spd > 0)

p.wr2 <- plot.windrose(data = wind_daily_renamed,
              spd = "spd",
              dir = "dir")

p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = 2) +
  theme(strip.text.x = element_text(size = 25)) + theme(panel.spacing = unit(2, "lines"))
p.wr3

################3

# load("Data/UI_angle.RData")
# 
# wind_plot_prep <- UI_angle %>% 
#   select(-t, -ui.saws)
# 
# wind_plot <- wind_plot_prep %>% 
#   as_tibble(rownames = "site") %>% 
#   group_by(site) %>% 
#   mutate_at(vars(-site), rescale) %>% 
#   tail(4) 
# 
# ggradar(wind_plot)
# 
# ??countmax

```

# ANOVA
- Relationship between duration, year and distance as a function of the mean intensity.

```{r}
options(scipen = 999)
anova_func <- function(df){
  sites_aov <- aov(intensity_mean ~ site * duration * distance, data = df)
return(sites_aov)
}

OISST_anov <- anova_func(df = OISST_upwell_base)
summary(OISST_anov)
CMC_anov <- anova_func(df = CMC_upwell_base)
summary(CMC_anov)

SACTN_anov <- sites_aov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
summary(SACTN_anov)
```

Plot showing upwelling at the different sites for each of the different distances. Scatter plot

```{r}
library(ggpubr)
year_func <- function(df){
  Filtered <- df %>%
filter(year(date_start) %in% seq(2010, 2016, 1)) 
}

CMC_filtered <- year_func(df = CMC_upwell_base)
OISST_filtered <- year_func(df = OISST_upwell_base)
SACTN_filtered <- year_func(df = SACTN_upwell_base)

CMC_SEAPOINT <- CMC_filtered %>% 
  filter(site == "Sea Point")
CMC_SB <- CMC_filtered %>% 
  filter(site == "Saldanha Bay")
CMC_PN <- CMC_filtered %>% 
  filter(site == "Port Nolloth")
CMC_LB <- CMC_filtered %>% 
  filter(site == "Lamberts Bay")

plot_scat_func <- function(df){
  ggplot(data = df, aes(x = date_start, y = intensity_mean)) +
  geom_point() +
  scale_x_date(expand = c(0, 0)) +
  facet_wrap(~distance, ncol = 2) +
  theme_grey() +
  labs(y = "Mean Intensity", x = NULL)
}
(CMC_SP_plot <- plot_scat_func(df = CMC_SEAPOINT))
(CMC_SB_plot <- plot_scat_func(df = CMC_SB))
(CMC_PN_plot <- plot_scat_func(df = CMC_PN))
(CMC_LB_plot <- plot_scat_func(df = CMC_LB))

combined_CMC <- ggarrange(CMC_SP_plot, CMC_SB_plot, CMC_PN_plot,CMC_LB_plot)
###################
## OISST
OISST_SEAPOINT <- OISST_filtered %>% 
  filter(site == "Sea Point")
OISST_SB <- OISST_filtered %>% 
  filter(site == "Saldanha Bay")
OISST_PN <- OISST_filtered %>% 
  filter(site == "Port Nolloth")
OISST_LB <- OISST_filtered %>% 
  filter(site == "Lamberts Bay")

plot_scat_func <- function(df){
  ggplot(data = df, aes(x = date_start, y = intensity_mean)) +
  geom_point() +
  scale_x_date(expand = c(0, 0)) +
  facet_wrap(~distance, ncol = 2) +
  theme_grey() +
  labs(y = "Mean Intensity", x = NULL)
}
(OISST_SP_plot <- plot_scat_func(df = OISST_SEAPOINT))
(OISST_SB_plot <- plot_scat_func(df = OISST_SB))
(OISST_PN_plot <- plot_scat_func(df = OISST_PN))
(OISST_LB_plot <- plot_scat_func(df = OISST_LB))
combined_OISST <- ggarrange(OISST_SP_plot, OISST_SB_plot, OISST_PN_plot,OISST_LB_plot)

############## 
# SACTN

plot_scat_func <- function(df){
  ggplot(data = df, aes(x = date_start, y = intensity_mean)) +
  geom_point() +
  scale_x_date(expand = c(0, 0)) +
  facet_wrap(~site, ncol = 2) +
  theme_grey() +
  labs(y = "Mean Intensity", x = NULL)
}
(SACTN_plot <- plot_scat_func(df = SACTN_filtered))

```

# Lolli plot

```{r}
plot_loli_func <- function(df){
  ggplot(df, aes(x = date_peak, y = duration)) + 
  geom_lolli(aes(colour = intensity_cumulative)) + 
  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
  facet_wrap(~distance, ncol = 2) +
  xlab("Date") + ylab("Event duration [days]") 
}

(OISST_SP_pl <- plot_loli_func(df = OISST_SEAPOINT))
(OISST_SB_pl <- plot_loli_func(df = OISST_SB))
(OISST_PN_pl <- plot_loli_func(df = OISST_PN))
(OISST_LB_pl <- plot_loli_func(df = OISST_LB))
combined_OISST_pl <- ggarrange(OISST_SP_pl, OISST_SB_pl, OISST_PN_pl,OISST_LB_pl)

#######
# CMC

(CMC_SP_pl <- plot_loli_func(df = CMC_SEAPOINT))
(CMC_SB_pl <- plot_loli_func(df = CMC_SB))
(CMC_PN_pl <- plot_loli_func(df = CMC_PN))
(CMC_LB_pl <- plot_loli_func(df = CMC_LB))
combined_CMC_pl <- ggarrange(CMC_SP_pl, CMC_SB_pl, CMC_PN_pl,CMC_LB_pl)

plot_loli_func <- function(df){
  ggplot(df, aes(x = date_peak, y = duration)) + 
  geom_lolli(aes(colour = intensity_cumulative)) + 
  scale_color_distiller(palette = "Spectral", name = "Cumulative \nintensity") + 
  facet_wrap(~site, ncol = 2) +
  xlab("Date") + ylab("Event duration [days]") 
}
(SACTN_plot <- plot_loli_func(df = SACTN_filtered))
```