### 2019-10-26 G1SST and MUR extraction.Rmd
UPDATE

---
title: "G1SST and MUR"
author: "Amieroh Abrahams"
date: "26 October 2019"
output: html_document
---

This scripts extracts the MUR and G1sst data

```{r}
bbox <- data.frame(BC = c(-35, -25, 15, 20)) # Benguela Current
library(stringr)
library(tidyverse)
library(reshape2)
library(ncdf4) # library for processing netCDFs
library(plyr)
library(lubridate)
library(data.table)
library(doMC); doMC::registerDoMC(cores = 7)
ncDir <- "/home/amieroh/Documents/Data/Datasets/G1SST/daily/2010"
csvDir <- "/home/amieroh/Documents/tester"

# #          1         2         3         4
# # 12345678901234567890123456789012345678901
# # 20020601-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc
#          1         2         3         4         5         6
# 1234567890123456789012345678901234567890123456789012345678901
# 20100609-JPL_OUROCEAN-L4UHfnd-GLOB-v01-fv01_0-G1SST_subset.nc

# ncFile <- '/home/amieroh/Documents/Data/Datasets/MUR/daily/20020606-JPL-L4UHfnd-GLOB-v01-fv04-MUR.nc'
  ncFun <- function(ncFile = ncFile, region = region, csvDir = csvDir) {
    coords <- bbox[, region]
    nc <- nc_open(ncFile)
    pathLen <- nchar(paste0(ncDir, "/")) + 1
    fNameStem <-
      substr(basename(ncFile), 10, 58)
    fDate <- substr(basename(ncFile), 1, 8)
    LatIdx <- which(nc$dim$lat$vals > coords[1] & nc$dim$lat$vals < coords[2])
    LonIdx <- which(nc$dim$lon$vals > coords[3] & nc$dim$lon$vals < coords[4])
    sst <- ncvar_get(nc, varid = "analysed_sst") %>%
      round(4)
    dimnames(sst) <- list(lon = nc$dim$lon$vals,
                          lat = nc$dim$lat$vals)
    nc_close(nc)
    sst <- as_tibble(melt(sst, value.name = "temp"))
    sst$t <- ymd(fDate)
    na.omit(sst)
    fwrite(sst,
           file = paste0(csvDir, "/", region, "-", fNameStem, "-", strtDate, "-", endDate, ".csv"),
           append = TRUE, col.names = FALSE)
    rm(sst)
  }

  ncList <- list.files(path = paste0(ncDir), pattern = "*.nc", full.names = TRUE, include.dirs = TRUE)
  ncFirst <- head(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
  ncLast <- tail(list.files(path = paste0(ncDir, "/"), pattern = "*.nc", full.names = FALSE), 1)
  strtDate <- str_sub(ncFirst, start = 1, end = 8)
  endDate <- str_sub(ncLast, start = 1, end = 8)

llply(ncList, ncFun, region = "BC", csvDir = csvDir, .parallel = TRUE)
```

```{r}
G1SST_prod <- G1SST %>%
  select(lon, lat) %>%
  unique() %>%
  mutate(product = "G1SST")

MUR_prod <- MUR %>%
  select(lon, lat) %>%
  mutate(product = "MUR")

sat_pixels <- sat_data %>%
select(product, lon, lat) %>%
unique()

match_func <- function(df){
  df <- df %>%
    dplyr::rename(lon_site = lon, lat_site = lat)
  G1SST_index <- G1SST_prod[as.vector(knnx.index(as.matrix(G1SST_prod[,c("lon", "lat")]),
                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
    cbind(., df)
  MUR_index <- MUR_prod[as.vector(knnx.index(as.matrix(MUR_prod[,c("lon", "lat")]),
                                             as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>%
    cbind(., df)
  res <- rbind(G1SST_index, MUR_index)
  return(res)
}

pixel_match <- site_pixels %>%
  group_by(site) %>%
  group_modify(~match_func(.x))

G1SST_fill <- right_join(G1SST, filter(pixel_match, product == "G1SST"), by = c("lon", "lat"))
MUR_fill <- right_join(MUR, filter(pixel_match, product == "MUR"), by = c("lon", "lat"))

# Clean up some RAM space
rm(MUR, G1SST); gc()

selected_sites <- c("Port Nolloth", "Lamberts Bay", "Sea Point", "Saldanha Bay")

G1SST_fill <- G1SST_fill %>%
  filter(site %in% selected_sites)

MUR_fill <- MUR_fill %>%
  filter(site %in% selected_sites)

load("Data/UI_angle.RData")
upwelling <- UI_angle %>%
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t),
  #        max_t = max(t)) %>%
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>%
  select(-data, -clim) %>%
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance)

detect_event_custom <- function(df){
  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
  return(res)
}

ts2clm_custom <- function(df){
  # The climatology base period used here is up for debate...
  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
  return(res)
}

# Calculate the upwelling event metrics
upwelling_detect_event <- function(df){
  upwell_base <- df %>%
    dplyr::rename(t = date) %>%
    group_by(site, product, heading, distance, lon, lat) %>%
    group_modify(~ts2clm_custom(.x)) %>%
    left_join(upwelling, by = c("site", "t")) %>%
    filter(!is.na(exceedance)) %>%
    group_by(site, product, heading, distance, lon, lat) %>%
    group_modify(~detect_event_custom(.x))
  }

G1SST_upwell_base <- upwelling_detect_event(df = G1SST_fill)
save(G1SST_upwell_base, file = "Data/G1SST_upwell_base.RData")
MUR_upwell_base <- upwelling_detect_event(df = MUR_fill)
save(MUR_upwell_base, file = "Data/MUR_upwell_base.RData")
```