### 2020-09-30 site_map.Rmd
update

---
title: "site_map"
author: "Amieroh Abrahams"
date: "01 August 2019"
output: html_document
---

# Loading Libraries

```{r prelim_opts, echo=FALSE}
knitr::opts_chunk$set(
  comment = "R>",
  warning = FALSE,
  message = FALSE 
)

library(tidyverse)
# library(plyr)
library(lubridate)
library(zoo)
library(FNN)
library(data.table)
library(viridis)
library(ggrepel)
library(plyr)
library(maptools)
library(sp)
library(geosphere)
library(PBSmapping)
library(scales)
library(grid)
library(gridExtra)
library(circular)
library(fossil)
library(mapproj)
library(reshape2)
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)
library(ggpubr); theme_set(theme_pubr())# library(devtools)
# install_github("marchtaylor/sinkr")

## Functions
source("Functions/earthdist.R")
source("Functions/wind.rose.R")
source("Functions/theme.R")
source("Functions/scale.bar.func.R")
```

```{r}
load("Data/site_list_sub.Rdata")
load("Data/africa_coast.RData")
xtable(site_list_sub, auto = TRUE)
west <- site_list_sub[-5,]
west$coast <- "west" # Chnages wc to west

site_list_sub <- site_list_sub[-5,]

load("Data/sa_bathy.RData")
## Downloading the bathy data from NOAA
# Download mid-res bathymetry data
# sa_lat <- c(-38, -24.5); sa_lon <- c(11.5, 35.5)
# sa_bathy <- as.xyz(getNOAA.bathy(lon1 = sa_lon[1], lon2 = sa_lon[2], lat1 = sa_lat[1], lat2 = sa_lat[2], resolution = 4))
# colnames(sa_bathy) <- c("lon", "lat", "depth")
# sa_bathy <- sa_bathy[sa_bathy$depth <= 0,]
# save(sa_bathy, file = "Data_P1/bathy/sa_bathy.RData")

# Loading in the newly downloaded bathymetry data
load("Data/sa_bathy.RData")
load("Data/africa_coast.RData")

# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
shore.normal.transect <- function(site, width = 2){
  # Find the site on the coastline and it's nearest neighbour points
  coords <- data.frame(lon = site$lon, lat = site$lat)
  coords2 <- knnx.index(africa_coast[,1:2], as.matrix(coords), k = 1)
  coords3 <- data.frame(site = site$site, africa_coast[c(coords2-width, coords2+width),]) 
  coords3 <- coords3[2:1,1:3]
  # Define the shore normal transect bearing
  heading <- earth.bear(coords3[1,2], coords3[1,3], coords3[2,2], coords3[2,3]) + 90
  if(heading >= 360){
    heading <- heading-360
  } else {
    heading <- heading
  }
  heading2 <- data.frame(site = site$site, lon = site$lon, lat = site$lat, heading)
  return(heading2)
}

# Creating the transects
site_transects <- data.frame()
for(i in 1:length(west$site)){
 site <- west[i,]
 site_transect <- shore.normal.transect(site, 2)
 site_transects <- rbind(site_transects, site_transect)
}

# Manually correcting Sea Point and Kommetjie
site_transects$heading[4] <- 290 
# save(site_transects, file = "Data/site_transects.RData")
load("Data/site_transects.RData")

# This function takes one site (e.g. one set of lon/lats) and calculates a shore normal transect
# It then extracts a lat/ lon point every X kilometres until reaching a specified isobath

transect.pixel <- function(site, distances){
  # Extract coordinates
  coords <- data.frame(lon = site$lon, lat = site$lat)
  # Find lon/ lats every X metres 
  pixels <- data.frame()
  # deep <- 999
  # distance_multiplier <- 1
  # while(deep > isobath){
  for(i in 1:length(distances)){
    coords2 <- as.data.frame(destPoint(p = coords, b = site$heading, d = distances[i]))
    sitesIdx <- knnx.index(sa_bathy[,1:2], as.matrix(coords2), k = 1)
    bathy2 <- sa_bathy[sitesIdx,]
    bathy2 <- bathy2[complete.cases(bathy2$depth),]
    bathy3 <- data.frame(site = site$site, lon = bathy2$lon, lat = bathy2$lat, 
                         heading = site$heading, 
                         distance = distances[i])
    pixels <- rbind(pixels, bathy3)
    coords <- coords2
  }
  if(nrow(pixels) < 1){
    pixels <- data.frame(site, depth = NA)
  }else{
    pixels <- pixels
  }
  return(pixels)
}

# Pixel points
site_pixels <- data.frame()
for(i in 1:length(west$site)){
  site <- site_transects[i,]
  #site_pixel <- transect.pixel(site, c(10000, 20000, 30000, 40000, 50000)) # RWS: fixed error
  site_pixel <- transect.pixel(site, c(10000, 25000, 50000))
  site_pixels <- rbind(site_pixels, site_pixel)
}

# New site pixels [AA,changesd some time recorded the land temp instad]
# save(site_pixels, file = "Data/site_pixels_new.RData")
# load("Data/site_pixels_new.RData")
# save(site_pixels, file = "Data/site_pixels.RData")
 load("Data/site_pixels.RData")

# Bounding box
  # Only one is made in order to know how large the the geom_point() squares should be made to match
bbox <- data.frame(xmin = destPoint(p = site_pixels[1,2:3], b = 270, d = 12500)[1],
                   xmax = destPoint(p = site_pixels[1,2:3], b = 90, d = 12500)[1],
                   ymin = destPoint(p = site_pixels[1,2:3], b = 180, d = 12500)[2],
                   ymax = destPoint(p = site_pixels[1,2:3], b = 0, d = 12500)[2])

## Determining the temperature at the various distances from the coast
```

```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Map representing the study region along the South African coastline. The black points represent the location of their *in situ* temperatures and approximations of the pixels used along the shore normal trasect from the satellite sea surface temperatures shown with black boxes", fig.height=6, fig.width=10, fig.pos="H"}

load("Data/south_africa_coast.RData")
load("Data/africa_coast.RData")
load("Data/sa_provinces_new.RData")
load("Data/site_list_sub.Rdata")
load("Data_coast_angle/site_pixels.RData")
load("~/Documents/HONOURSPROJECT-master/data/africa_coast.RData")
site_pixels <- read_csv("site_pixels.csv")

load("~/Documents/Upwelling_MCS/Data/sa_bathy.RData")

#load("Data/site_pixels.RData")
#load("MUR.RData")
names(south_africa_coast)[1] <- "lon"

# Manually divide up coastline
wc <- south_africa_coast[291:410,]

site_list_sub <- site_list_sub[-5,]

site_list_sub_map <- site_list_sub %>%
  dplyr::select(site,lat,lon)

# Setting up the theme
theme_set(theme_bw())
limits <- c(12,28) # for colour bar
#breaks <- seq(6, 30, 2) # Create breaks to be used for colour bar

# Define plotting parameters
sa_lats <- c(-37, -27); sa_lons <- c(14, 25)

site_map <- ggplot(data = south_africa_coast, aes(x = lon, y = lat)) + 
   geom_polygon(aes(group = group),fill = "grey", colour = "grey", size = 0.1, show.legend = FALSE) +
  stat_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth), 
               colour = "black", size = 0.2, binwidth = 200, na.rm = TRUE, show_guide = FALSE)+
  geom_point(data = site_list_sub, aes(x = lon, y = lat), alpha = 0.8, size = 2) +
  geom_point(data = site_pixels, aes(x = lon, y = lat, fill = "red"), colour = "red", shape = 15, alpha = 0.8, size = 2.1) +
  geom_label_repel(data =site_list_sub, aes(x = lon, y = lat, label = site),
                     size = 4, box.padding = 2, nudge_x = 0.7, nudge_y = 0.2,
                     segment.alpha = 0.4, force = 0.1, font = "Palatino") +
  xlab("") + ylab("") +
    # annotate("text", label = "ATLANTIC\nOCEAN", x = 16.5, y = -31.5,
    #          size =5, angle = 0, colour = "black") +
  annotate("text", label = "Cape Columbine", x = 16.0, y = -30.5,
             size =5, angle = 45, colour = "black")+
  annotate("text", label = "Cape Peninsula", x = 16.5, y = -33.9,
             size =5, angle = 45, colour = "black")+
  guides(fill = guide_colourbar()) +
  scale_y_continuous(breaks = seq(-34, -28, by = 2),
                     labels  = c("34°S","32°S","30°S","28°S")) +
  scale_x_continuous(breaks = seq(16, 20, by = 2),
                      labels = c("16°E", "18°E", "20°E")) +
 coord_fixed(ratio = 1, xlim = c(15.4, 20), ylim = c(-34.4,-27),
    expand = TRUE) +
  theme(axis.text = element_text(size = rel(1.2), colour = "black", family = "Palatino"),
        axis.text.x = element_text(vjust = 1, family = "Palatino", size = 22),
        axis.text.y = element_text(hjust = 1, family = "Palatino", size = 22),
        axis.title.x = element_text(vjust = 0, family = "Palatino"),
        axis.title.y = element_text(angle = 90, vjust = 0.3, family = "Palatino"),
        axis.ticks = element_line(colour = "black"),
        axis.ticks.length = unit(0.4, "cm"),
        #axis.ticks.length = unit(0.15, "cm"),
        axis.ticks.margin = unit(0.1, "cm"),
        panel.background = element_rect(fill = "white", colour = "black", size = 0.5),
        #panel.border = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, size=2),
        #(colour =" black"),
        panel.grid.major.x = element_blank(),
        #panel.grid.major.y = element_line(linetype = "dotted", size = 0.2, colour = "black"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        #panel.margin = unit(0.25, "lines"),
        strip.text = element_text(size = rel(1.0), family = "Palatino"),
        strip.background = element_rect(fill = "white", colour = NA),
        strip.text.x = element_text(),
        strip.text.y = element_text(angle = -90),        
        plot.background = element_rect(colour = "white"),
        plot.title = element_text(size = rel(1.2), vjust = 1),
        plot.margin = unit(c(1, 1, 0.5, 0.5), "lines"))

site_map

africa_map <- ggplot(africa_coast, aes(x = lon, y = lat)) +
  theme_bw() +
  coord_equal() +
  geom_polygon(aes(group = group), colour = "black", fill = "grey80") +
  geom_rect(data=africa_coast, aes(xmin=15.0, xmax=21.0, ymin=-34.4, ymax=-29.5),  colour = "red", fill = NA, size = 0.35) +
  geom_polygon(data = sa_provinces_new, (aes(group = group))) +
  #annotate("text", label = "Africa", x = 16.0, y = 15.0, size = 3) +
  theme(panel.border = element_rect(colour = "black", size = 0.4),
        plot.background = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  coord_map(xlim = c(-20, 53), ylim = c(-36, 38), projection = "mercator")
africa_map

sa_4 <- site_map +
annotation_custom(grob = ggplotGrob(africa_map),
xmin = 17, xmax = 21.5,
ymin = -28.8, ymax = -26.6)
sa_4
```

# SIte map 2

```{r}

# site_list_sub_map <- site_list_sub %>% 
#   select(site,lat,lon)
# 
# mapping_plot <- function(df, xy = "salmon"){
#   mapping <- df %>% 
#     ggplot(aes(x = lon, y = lat)) +
#     geom_polygon(data = south_africa_coast, aes(group = group), fill = "white") +
#     coord_cartesian(xlim = c(12, 36), ylim = c(-32, -20)) +
#     geom_path(data = sa_provinces_new, aes(group = group)) +
#     geom_point(data = df, colour = xy, size = 4)  +
#     geom_label_repel(data =df, aes(x = lon, y = lat, label = site),   
#                      size = 6, box.padding = 2, nudge_x = -0.5, nudge_y = 0.2, 
#                      segment.alpha = 0.4, force = 0.1) +
#     coord_equal(xlim = c(15, 35), ylim = c(-37, -27)) +
#     annotate("text", label = "INDIAN\nOCEAN", x = 35.00, y = -34.0, size = 6.0, angle = 0, colour = "black") +
#     annotate("text", label = "ATLANTIC\nOCEAN", x = 16.00, y = -37.0, size = 6.0, angle = 0, colour = "black") +
#     # geom_segment(aes(x = 17.2, y = -32.6, xend = 15.2, yend = -29.5),
#     #               arrow = arrow(length = unit(0.3, "cm")), size = 0.1, colour = "black") +
#     #   annotate("text", label = "Benguela", x = 14.0, y = -31.8, size = 3.5, angle = 298, colour = "black") +
#     #   geom_segment(aes(x = 33, y = -29.5, xend = 29.8, yend = -33.0),
#     #               arrow = arrow(length = unit(0.3, "cm")), size = 0.1, colour = "black") +
#     annotate("text", label = "Agulhas", x = 31.7, y = -31.7, size = 7.5, angle = 53, colour = "black") +
#     scale_x_continuous(labels = scales::unit_format(unit = "°E", sep = "")) +
#     scale_y_continuous(labels = scales::unit_format(unit = "°S", sep = "")) +
#     labs(x = NULL, y = NULL) +
#     coord_cartesian(xlim = c(10.5, 39.5), ylim = c(-39.5, -25.5), expand = F) +
#     scaleBar(lon = 32.0, lat = -38.7, distanceLon = 200, distanceLat = 50, 
#              distanceLegend = 90, dist.unit = "km", arrow.length = 100,
#              arrow.distance = 130, arrow.North.size = 3) +  
#     coord_equal() +
#     theme(aspect.ratio = 1)  +
#     theme_bw() +
#     theme(panel.border = element_rect(fill = NA, colour = "black", size = 1),
#           axis.text = element_text(colour = "black", size = 22),
#           axis.title = element_text(colour = "black", size = 22))
#   return(mapping)
# }
# 
# Map <- mapping_plot(df = site_list_sub_map, xy = "salmon")
```

# Bathy plot

```{r}
fig1_bathy <- ggplot() + theme_bw() + #coord_equal() + 
  geom_raster(data = sa_bathy, aes(x = lon, y = lat, fill = depth)) +
  #geom_raster(data = isobath200, aes(x = lon, y = lat, fill = coast)) +
  stat_contour(data = sa_bathy, aes(x = lon, y = lat, z = depth, alpha = ..level..), 
               colour = "black", size = 0.2, binwidth = 200, na.rm = TRUE, show_guide = FALSE) +
  geom_polygon(data = south_africa_coast, aes(x = lon, y = lat, group = group), 
               size = 0.1, colour = "black", fill = "grey80") +
  #geom_path(data = sa_provinces_new, aes(x = long, y = lat, group = group)) +
  # geom_polygon(data = scBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 
  #              colour = "green", fill = "green") +
  # geom_polygon(data = wcBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 
  #              colour = "blue", fill = "blue") +
  # geom_polygon(data = ecBox, aes(x = lon, y = lat, group = group), alpha = 0.20, 
  #              colour = "orange", fill = "orange") +
  geom_point(data = site_pixels, aes(x = lon, y = lat, colour = site), alpha = 0.8, size = 3) +
  geom_point(data = site_pixels, aes(x = lon, y = lat), colour = "red", shape = 0, alpha = 0.8, size = 2.1) +
  #geom_rect(data = test, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
            #alpha = 1, colour = "red", size = 0.1, linetype = 1) +
   scale_y_continuous(breaks = seq(-34, -30, by = 2),
                     labels  = c("34°S","32°S","30°S")) +
  scale_x_continuous(breaks = seq(16, 18, by = 2),
                      labels = c("16°E", "18°E"))+
  scale_alpha_continuous(breaks = c(-200, -1000, -2000, -3000, -4000, -5000),
                         guide_legend(title = "depth (m)")) +
  scale_fill_gradient(low = "steelblue4", high = "steelblue1", na.value = "steelblue4", 
                      breaks = c(-1000, -2000, -3000, -4000, -5000),
                      guide_legend(title = "depth (m)")) +
  labs(title = NULL, x = NULL, y = NULL) +
  guides(colour = guide_legend(override.aes = list(size = 3))) +
  #guide_legend(ncol = 2) +
  ### Annotate specific things:
  annotate("text", label = "Cape \n Agulhas", x = 20.0, y = -35.1, size = 2., colour = "white") +
  #annotate("text", label = "False Bay", x = 18.66, y = -34.55, size = 2.2) +
  annotate("text", label = "Algoa \n Bay", x = 26.4, y = -34.0, size = 2., colour = "white") +
  annotate("text", label = "Cape \n Point", x = 18.0, y = -34.4, size = 2, colour = "white") +
  annotate("text", label = "Hamburg", x = 28.3, y = -33.28611111, size = 2, colour = "white") +
  theme(#panel.background = element_rect(fill = "steelblue4", colour = NA),
        #panel.border = element_blank(),
        #axis.title = element_blank(),
        #axis.text = element_blank(),
        #axis.ticks = element_blank(),
        #panel.grid.minor = element_blank(),
        #panel.grid.major = element_blank(),
        #plot.title = element_blank(),
        #legend.text = element_text(size = 7, colour = "White"),
        #legend.title = element_text(size = 7, colour = "White"),
        legend.key = element_rect(colour = NA, size = 0.2),
        legend.key.height = unit(0.4, "cm"),
        legend.background = element_blank(),
        legend.justification = c(1,0), legend.position = c(0.5, 0.4)) +
  coord_cartesian(xlim = sa_lons, ylim = sa_lats)
```